---
title: "Lab 5"
author: "Corinna Hong"
date: "February 6, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r, load packages}

library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)

```


## Monthly US Energy Consumption (Renewables)
```{r, read in data}

us_renew <- read_csv(here("data","renewables_cons_prod.csv")) %>% 
  clean_names()

```


- Make descriptions all lowercase
- Only keep observations where "description" variable contains "consumption"
- Remove any observations where "description" variable contains "total"

```{r}

renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>%
  filter(!str_detect(description, pattern = "total"))

```

- convert 'yyyymm' column to a date

```{r}

renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% # Pulls just year and month from yr_mo_day column. Stores in way feasts and fable will understand
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value) # draop NAs in these two columns

# Version where month and year are in separate columns

renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% # label = true gives months the 3 letter label
  mutate(year = year(yr_mo_day))

```

```{r}

renew_gg <- ggplot(data = renew_date, aes(x = month_sep, y = value, group = description)) +
  geom_line(aes(color = description)) # using aes(color = ) down here after using (group = ) above is redundant in terms of separating lines by description. This gives it color tho!

renew_gg
  
```

Updating colors with paletteer palettes

```{r}

# use view(palette_d_names) to see all of them

renew_gg +
  scale_color_paletteer_d("ggsci::planetexpress_futurama") +
  theme_classic()

```

### Coerce renew_parsed to a tibble

```{r}

renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)

```

Look at time series data in a couple of different ways
```{r}

renew_ts %>% autoplot(value)

renew_ts %>% gg_subseries(value)

# renew_ts %>% gg_season(value)

ggplot(data = renew_parsed, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year)) +
  facet_wrap(~description, ncol = 1, scales = "free", strip.position = "right")


```

### Just look at hydroelectric 

```{r}

hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)

ggplot(hydro_ts, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year))

```

### What if I want quarterly consumption

```{r}

hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarize(avg_consumption = mean(value))

head(hydro_quarterly)

```

### Decompose hydro_ts

```{r}

dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))


components(dcmp) %>%  autoplot()

```

```{r}

# hist(components(dcmp))

```

Now let's look at the ACF.

```{r}

hydro_ts %>% 
  ACF(value) %>% 
  autoplot()

```

# DANGER

```{r}

hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))

```


### MAke a world map

```{r}

# world <- read_sf(dsn = here("data", "TM_WORLD_BOARDERS_SIMPL-0.3-1"), layer = "TM_WORLD_BOARDERS_SIMPL-0.3")

```

